---
title: "Probability and Statistics Assignment 1"
author: "Harshad Kumar Elangovan - 19200349"
date: "01/10/2019"
output: html_document
---
<br/>
<br/>
The table with exam marks of each student is created using the below commands.
```{r}
studentname=c("John","Anne","Terry","Fred","Maria")
exam1=c(92,75,98,62,79)
exam2=c(82,96,60,55,72)
results=data.frame(studentname,exam1,exam2)
print(results)
```

### <u>Part A</u> 
#### What code creates the list of exam2 marks?

The code for creating the list of exam2 marks is

```{r}
exam2=c(82,96,60,55,72)
```
<br/>
<br/>

### <u>Part B</u>
#### Who has a better grade, Maria or Terry?

From the table, we can see that Maria and Terry have the same grade 'B'. But the average of Terry is more than that of Maria. So, **Terry has better grade than Maria**.

```{r}
results$avg = ((results$exam1 + results$exam2)/2)
results$grade = cut(results$avg, breaks = c(0,49,69,84,100),labels=c("D","C","B","A"))
results = results[order(-results$avg),]
print(results)
```

### <u>Part C</u>
#### Suppose you wanted to print out only the names and exam 1 marks of the students. What code would you type?

We can display the names and exam 1 of the students separately by creating another dataframe with only those two columns in it. This can be done using the below code,

```{r}
results1 = data.frame(results$studentname,results$exam1)
print(results1)
```

### <u>Part D</u>
#### Mention the mean and median of Pulse1 and Pulse2 and which do you think is pulse after exercise?

```{r}
pulse = read.table("C:\\Users\\harshie\\Documents\\UCD\\Probability\\Lab1\\pulse.txt", header=TRUE)
pulse = data.frame(pulse)
```

```{r}
summary(pulse$pulse1)
summary(pulse$pulse2)
```

From the above table, the mean and median of both the pulses are as follows,

<u>Pulse1</u>

Mean:72.87

Median:71.00
<br/>
<br/>

<u>Pulse2</u>

Mean:80

Median:76

The pulse 2 with mean and median value higher is the pulse after exercise.

### <u>Part E</u>
#### Mention the mean and median of Ran = 1 and Ran = 2 and Which group do you think ran and which walked?

The mean and median of the pulses with Ran value 1 and 2 can be found using the below mentioned function in markdown. This will produce the output based on the conditions.

```{r}
summary(pulse$pulse2[pulse$ran ==1])
summary(pulse$pulse2[pulse$ran ==2])
```
From the above table, the mean and median of both the pulses are as follows,

<u>Ran = 1</u>

Mean:92.51

Median:88.00
<br/>
<br/>

<u>Ran = 2</u>

Mean:72.32

Median:70.00

So, from the above mean and median values, the pulse with Ran =1 ran while exercising and pulse with Ran=2 walked.

### <u>Part F</u>
#### What is the most common (modal) pulse1 measurement?

From the output below, the most common measurement in pulse1 is 68.

```{r}
table(pulse$pulse1)
```

#### What is the most common (modal) pulse2 measurement?

From the output below, the most common measurement in pulse2 is 76.

```{r}
table(pulse$pulse2)
```

### <u>Part G</u>
#### Sketch the histogram of the weights of female volunteers here:

The code for producing the histogram of the weights of female volunteers is as follows,

```{r}
hist(pulse$weight[pulse$sex==2],xlab="Weight",main="Histogram of Female Volunteers based on weight")
```