---
title: "Adv Data Prog with R_Final Project STAT40830_19200324"
author: "GUL RUKH AHMAD"
date: "16/08/2020"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Question 1**

The dataset is loaded as tibble with the help of rio library.

Also Columns 1, 16, 17, 18 and 25 are loaded as characters

Columns 2 and 14 are converted to factors

Column 15 is loaded as integers and rest other columns are doubles.

There is 3659 records loaded in exo dataset

```{r}
set.seed(19200324) # Setting seed

#install.packages("rio")
library(rio) # rio package is installed and library rio is called which helps in importing file directly as tibble
exo <- import("exo_data.csv", setclass = "tibble")
library(dplyr)
library(magrittr)
# Converting the two variables flag (Column2) and meth (Column14) as factors
exo<-exo %>% mutate(flag=factor(flag)) %>% mutate(meth=factor(meth))
head(exo) # Checking first few records from the top
glimpse(exo) # Checking every column in the tibble
```

**Question 2 **

Unknown method of discovery is removed.

We have 3596 records in exo1 dataset.

```{r}

# Checking the levels of method of discovery (meth) column
# It has six levels imaging, microlensing, RV, timing, transit and a blank level
levels(exo$meth)
# We need to remove all the records associated with blank level for  method of discovery.
# In other words we need to keep only five known levels(known method of discovery) and omit unknown method of discovery
# We have filtered records having these five levels and dropped the unused blank level as well.
# New dataset exo1 is created 
exo1<-exo %>% filter(meth=='imaging'| meth=='microlensing' | meth=='RV'| meth=='timing' | meth=='transit') %>%droplevels(exo$meth)
                  
# Checking the levels of meth column in exo1 dataset which is as per requirement (the five known method of discovery)
levels(exo1$meth)
glimpse(exo1)
```

**Question3 **

We have to show the relationship between log-distances from the sun and method of discovery.

Log-distances from the sun is continuous variable and method of discovery is discrete categorical variable.

We will draw box-plot in such case to show the relationship.

Observation from Box Plot-:

1. There are no outliers for imaging method of discovery. Rest other methods of discovery has outliers

2. Microlensing and RV methods of discovery are normally distributed.

3. Imaging method of discovery is left-skewed (negatively skewed)

4. Timing and Transit method of discovery is slightly right-skewed (positively skewed)


```{r}

# Box plot is drawn to show the relationship

library(ggplot2)
r <- ggplot(data = exo1, aes(x=meth,y=log(dist),fill=meth))+
  geom_boxplot(na.rm=TRUE)
r+
  labs(
    title = "Box plot for relationship between log-distance from Sun and method of discovery",
    x = 'method of discovery',
    y = 'log of distance from sun'
    
  )
```



**Question 4**

Scatter plot as per the given requirement is drawn between log mass and log distances separated by method of discovery

In fact we have drawn two plots for this-:

One plot has all the records of scatter plot in a single grid separated by different colors for method of discovery

For Other plot we have just added facet_wrap to make the plots in different grid having different color representation
for the  methods of discovery



```{r}

# Facet_wrap is used here
#install.packages("ggiraph")
library(ggiraph)
library(viridis)
# onclick column is added in exo1 dataset with values as such which will meet the requirement of 
# exoplanet's page being opened when cursor is clicked on the scatterplot 
exo1$onclick <- sprintf("window.open(\"%s%s\")",
                        "http://www.openexoplanetcatalogue.com/planet/",
                        exo1$id)
p<-ggplot(exo1, aes(x = log(dist), y = log(mass), colour = meth))+
  
  xlab('log of Mass')+
  ylab('log of Distance')+
  ggtitle('Scatter Plot for log mass versus log distance')+
  
  scale_color_discrete(name="Methods of Discovery")+ facet_wrap(~meth) + theme_light()
p1 <- p + geom_point_interactive(aes(tooltip = id,
                                     data_id = id,onclick = onclick),
                                 size = 2,na.rm=TRUE)
ggiraph(code = print(p1), width = 0.9)


# Facet_wrap is not used here
q<-ggplot(exo1, aes(x = log(dist), y = log(mass), colour = meth))+
  
  xlab('log of Mass')+
  ylab('log of Distance')+
  ggtitle('Scatter Plot for log mass versus log distance')+
  
  scale_color_discrete(name="Methods of Discovery")+ theme_light()
q1 <- q + geom_point_interactive(aes(tooltip = id,
                                     data_id = id,onclick = onclick),
                                 size = 2,na.rm=TRUE)
ggiraph(code = print(q1), width = 0.9)

```


**Question 5**

We have renamed radius column to jupiter_radius and created new column earth_radius which is 11.2 times the 
jupiter_radius


```{r}

# radius is renamed to jupiter radius and new column earth radius is created which is 11.2 times the jupiter radius
exo1<-exo1 %>% rename(jupiter_radius=radius) %>% mutate (earth_radius=11.2*jupiter_radius)
glimpse(exo1) # checking all the columns of dataset now

```

**Question 6**

There are 2732 records in exo2 dataset after exclusion of missing values from log-earth radius and log -period column

After kmeans is performed we get following number of records in each cluster-:

Cluster1 = 441
Cluster2 = 751
Cluster3 = 384
Cluster4 = 1156


```{r}
# Keeping only those rows in exo2 dataset where log-earth radius and log-period has no missing values
exo2<-exo1 %>% filter(!is.na(log(earth_radius)) & !is.na(log(period)) )
glimpse(exo2)
# In exo3 dataset we are keeping only earth radius and period variable on which kmeans is to performed 
exo3<-exo2 %>%  select(earth_radius,period)
# Performing kmeans on log-earth radius and log-period with four clusters
k_cluster<-kmeans(log(exo3),centers = 4)
# Records in each cluster is found out after kmeans is executed
k_cluster$size

```

**Question 7**

Cluster 1 having 441 records is assigned cold_gas_giants clustering label
Cluster 2 having 751 records is assigned others clustering label
Cluster 3 having 384 records is assigned hot_jupiters clustering label
Cluster 4 having 1156 records is assigned rocky clustering label

The clustering label is assigned based on visualization of plot given in the link in the question.

These all new clustering label is incorporated in newly created type variable in the exo2 dataset

Finally we have drawn scatter plot for log-earth radius vs log-period highlighting these clusters.


```{r}

type<-k_cluster$cluster # type variable is created which is having the clusters found
#install.packages("tidyverse")
library(tidyverse)
exo2<-exo2 %>% mutate(type=factor(type)) # Adding type variable to exo2 dataset and converting it to factor
# Re-coding the labels of type variable. The clustering label is applied based on visualizing the plot as given in the # https://en.wikipedia.org/wiki/Exoplanet#/media/File:ExoplanetPopulations-20170616.png plot 
exo2<- exo2 %>% mutate(type = recode(type, "1"="cold_gas_giants", "2"="others", "3"="hot_jupiters", "4"="rocky"))

levels(exo2$type) # Checking the clustering labels
table(exo2$type) # Checking whether each clustering label has correct number of records as per our visual representation of plot given

#Required Scatter plot is drawn highlighting the four clusters
ggplot(data=exo2, aes(x=log(period), y=log(earth_radius), color=type )) + 
  geom_point() +
  xlab('log of Period')+
  ylab('log of Earth_Radius')+
  ggtitle('Scatter plot of log of Period vs log of earth radius highlighting four clusters')

# Note-: with the help of "stat_ellipse" we can highlight this cluster much better way as well
# Stat ellipse is the extra line of code added here
t<-ggplot(data=exo2, aes(x=log(period), y=log(earth_radius), color=type )) + 
  geom_point() +
  stat_ellipse(type = "t",geom = "polygon",alpha = 0.4)+
  xlab('log of Period')+
  ylab('log of Earth_Radius')+
  ggtitle('Scatter plot  highlighting four clusters with stat_ellipse')
t

```


**Question 8**

Violin plot is drawn to illustrate the relationship between log-mass (continuous variable) and type (discrete categorical variable)

Observations from violin plot-:

1. There are no outliers for cold_gas_giants type. Rest other types have outliers

2. Rocky is having normally distributed data

3. cold_gas_giants and hot_jupiters data are  slightly left skewed (negatively skewed)

4. Others type data are slightly right skewed (positively skewed)


```{r}

n <- ggplot(data = exo2, aes(x=type,y=log(mass),fill=type))+
  geom_violin(na.rm=TRUE)
n+
  labs(
    title = "Violin plot for relationship between log-mass vs cluster type",
    x = 'cluster type',
    y = 'log of mass'
    
  )

```


**Question 9**

lubridate is used which makes it easier to work with date and time

r_asc values are converted to seconds and the new value is kept in  newly created variable r_asc_seconds
decl values are converted to seconds and the new value is kept in newly created decl_seconds variable

Note-: There are no records under microlensing and timing method in exo2 (having 2732 observations) dataset.
Hence while drawing the celestial map (scatter plot ) of exoplanets ; these methods having no records are not shown

```{r}

library(lubridate) # lubricate is called which makes it easier to work with dates and times

# Converting values in r_asc to seconds and putting them in new variable r_asc_seconds
exo2<-exo2 %>% mutate(r_asc=str_replace_all(r_asc," ",":")) %>% mutate(r_asc=hms(r_asc)) %>% mutate (r_asc_seconds=period_to_seconds(r_asc))


# Converting values in decl to seconds and putting them in new variable decl_seconds
exo2<-exo2 %>% mutate(decl=str_replace_all(decl," ",":")) %>% mutate(decl=hms(decl)) %>% mutate (decl_seconds=period_to_seconds(decl))

# Checking total number of records in meth column for each methods in exo2 dataset.
# We observe that there are no records under microlensing and timing method 
table(exo2$meth)

# Celestial map for the exoplanets drawn- Scatter plot is drawn for same
# Note-: Since there are no records under microlensing and timing method; these are not shown in the plot
a<-ggplot(exo2, aes(x = r_asc_seconds, y = decl_seconds, colour = meth))+
  geom_point(na.rm=TRUE)
a+
  labs(
    title = "Celestial map for the exoplanets",
    x = 'Right ascension in seconds',
    y = 'Declination in seconds'
    
  ) 
```


**Question 10**

We have used exo1 dataset from question2 having 3596 records for this question

exo1_new dataset is created where method and year is grouped and we have the count of each method for that year as well as cumulative count of method also available.

exo1_new dataset has NA values which resulted in error while drawing animated time series graph. Hence removed the NA values records from the dataset
The error thrown was-:
Error in sprintf(paste0("%0", nc, "i"), frame) : 
  invalid format '%0NA'; use format %f, %e, %g or %a for numeric objects
  
  
Finally animated time series is plotted illustrating the evolution over time of the total number of exoplanets discovered for each method up to that year

Note-: we have drawn animated times series with facet_wrap included where the plots of each method of discovery is in different grid. We can plot the same animated time series without facet_wrap as well.


```{r}
# We have used exo1 dataset from question2 having 3596 records for this question

# exo1_new dataset is created  where method and year is grouped and we having the count of each method for that year and cumulative count of method is also available.
exo1_new <- exo1 %>% group_by(meth, year) %>%  summarise(method_count = length(meth)) %>%
  mutate(method_count = cumsum(method_count))
# Checking the dataset exo1_new for any NA values 
View(exo1_new)
# Note-: Animated time series when attempted to create with NA values thrown error hence omitting records with NA values
exo1_new <- na.omit(exo1_new)
library(gganimate)
library(gifski)

# Animated time series is plotted 
anim <- ggplot(exo1_new, aes(x = year, y = method_count,color=meth)) +
  geom_line(aes(group = meth)) +
  geom_point(aes(group = seq_along(year),size=1)) + 
  facet_wrap(~meth)+
  transition_reveal(year) +
  labs(title='evolution over time total number of exoplanets discovered for each method',xlab='year',ylab='method_count')+
  enter_fade() +
  exit_fade()

animation <- animate(anim, renderer = gifski_renderer(loop = T))
animation

```


```{r ,message=FALSE,warning=FALSE}

library(shiny)

ui <- shinyUI(
    fluidPage(
      # Title of Application
    titlePanel("Discovery of Exoplanet"),

    
    sidebarLayout(
        sidebarPanel(
            sliderInput("year",
                        "Yearwise Discovery of Exoplanets:",
                        min = 2009,
                        max = 2018,
                        value = 2015),
            
            selectInput("type",
                        label = "Type of Exoplanets Discovered:",
                        choices = c('rocky', 'hot_jupiters', 'cold_gas_giants', 'others','All'),
                        selected = 'rocky')
        ),

       
        mainPanel(
           plotOutput("ScatterPlot")
        )
    )
  )
)

server <- function(input, output) {
    output$ScatterPlot <- renderPlot({
      
        y <- exo2
        y <- reactive({
                      if(input$type == 'All'){
                        exo2[(exo2$year <= input$year),] 
                      }
                      else {
                         exo2[((exo2$year <= input$year) & (exo2$type == input$type)),] 
                      }
                       
                      })
       
        ggplot(y(), aes(x= log(mass),y = log(dist), color = meth)) +
          geom_point() + labs(title='Exoplanets on Scatter Plot ',xlab='log ofmass',ylab='log of distance')
          
    })
}

shinyApp(ui = ui, server = server)

```

**Question 12**

Linear regression model  with the help of inbuilt lm function is applied to the data with log-period as response variable and logs of host_mass, host_temp and axis as explanatory variables

Note-: Intercept is also included in the regression model

```{r}

# Note we are using exo2 dataset from Question 6 for our analysis here.
# We are selecting the covariates given in the question for fitting multiple  linear regression model
exo_covariate<-exo2 %>%  select(period, host_mass, host_temp, axis)
exo_covariate_final <- na.omit(exo_covariate) # We are excluding those rows where there is at least one missing value
# Taking the log of each covariate
exo_covariate_final<- exo_covariate_final %>% mutate(period = log(period),host_mass = log(host_mass),
                              host_temp = log(host_temp), axis = log(axis)) 
# Fitting the Linear regression model with lm function
regression_model<-lm(period ~ host_mass+host_temp+axis, data = exo_covariate_final)
```



**Question 13**


Interpretation of the fitted model-:

 1. Estimated coefficient for Intercept is 7.2178453 which means that for 1 unit increase in Intercept value keeping the predictor variables host mass, host temp and axis constant; the response variable- period increases by 7.2178453 unit. There is a positive correlation between Intercept and Period
 
 2. Estimated coefficient for host mass is -0.2599925 which means that for 1 unit increase in host mass value keeping the predictor variables host temp , axis and Intercept constant; the response variable- period decreases by 0.2599925  unit. There is a negative correlation between host mass and period
 
3. Estimated coefficient for host temp is -0.1578013 which means that for 1 unit increase in host temp value keeping the predictor variables host mass , axis and Intercept constant; the response variable- period decreases by 0.1578013  unit. There is a negative correlation between host temp and period

4. Estimated coefficient for axis is 1.4830300 which means that for 1 unit increase in axis value keeping the predictor variables host mass , host temp and Intercept constant; the response variable- period increases by  1.4830300  unit. There is a positive correlation between axis and period

5. P-value for estimated coefficient of Intercept, host_mass, host_temp and axis is less than 0.05 for all these which means that the explanatory variables host mass , host temp and axis are statistically significant at 95% Confidence Interval and they all contribute in explaining the model and all these variables helps in predicting the response variable period

6. R-squared value  is 0.9795563 which means 97.95% of the model is explained by the  explanatory variables. R squared value helps in interpreting as to how well the regression model fits the observed data. High R squared means better fit and low R squared value means bad fit of model to the data.

7. Adjusted R-square value is 0.9794568 which is almost equal to  R-squared value .It compares the goodness-of-fit for regression models that contain differing numbers of independent variables. Here we can conclude that the explanatory variable are doing a very good job in explaining the model

8. We have also shown the 95% CI value for the estimated coefficient of Intercept, host_mass, host_temp,axis.

9. With the help of ggResidpanel we have drawn plots like residual plot, qqplot etc which helps in further interpretation of the fitted regression model.

From the QQ-plot we could see that it is almost normal with some outliers in the tail end

From the histogram plot of residuals (studentized) vs density; we could see the density plot is also normally distributed

From the residual plot of  Predicted values vs Residuals we could see that the most of  values are hovering around 0
with some outliers. The residuals are having  symmetrical pattern and almost have a constant spread throughout the range. It is homoskedastic in nature. There is no funnel shape in residual vs predicted values graph which otherwise would have indicated non-constant variance.

In the residual vs leverage plot as well there is no strong evidence for heteroskedasticity

**10. Overall we can conclude that the fitted linear regression model is a good model **



```{r}

# Overall Summary of the fitted model
summary(regression_model)

# We can also get the summaries one by one as well

regression_model$coefficients # Estimated coefficients of Intercept and predictor variable
confint(regression_model, level=0.95) # 95 % Confidence Interval of the estimated coefficients

summary(regression_model)$r.squared  # R-squared value of our model
summary(regression_model)$adj.r.squared # Adjusted R-squared value of our model

#install.packages("ggResidpanel")
# With the help of ggResidpanel we are drawing various plots like Residual Plot, Q-Q Plot etc which helps us in analyzing the fitted values of the regression model
library(ggResidpanel)
resid_panel(regression_model,plots = "all")
resid_xpanel(regression_model, yvar = "response")
```



**Question 11 and 14**

We have created the Interactive shiny plot as per the question requirement and embedded the same into R-markdown






