---
title: "Probability R Assignment 4"
author: "Harshad Kumar Elangovan - 19200349"
date: "25/11/2019"
output: html_document
---

## Importing Icecream Data

```{r importing}
icecreamdata<-read.table("C:\\Users\\harshie\\Documents\\UCD\\Probability\\Lab4\\icecream.txt",header=TRUE)
```
## Part A:

Plotting the sales ~ temperature for the given data:
```{r plottingtempSales}
plot(icecreamdata$temperature,icecreamdata$sales, main = "Sales VS Temperature",xlab = "Temperature", 
     ylab = "Sales")
```

### What sort of relationship exists between distance and damage?

The temperature and sales and linearly dependent in this model. So, as the temperature increases, the sales of increases also increases.


## Part B

```{r cor}
cor(icecreamdata$temperature,icecreamdata$sales)
```

### What is the value of the correlation coefficient?


The value of the correlation coefficient is 0.9722422.

## Part 3

```{r lsfit}
lsfit(icecreamdata$temperature,icecreamdata$sales)
```

The equation of the regression model is

Icecream_Sales = Beta0 + Beta1*Temperature;

Icecream_Sales = 5.5689655 + 0.7586207*Temperture;


we can predict the sales of icecream when the temperature is 42, by solving the equation with its value,

```{r temp}
sales<-5.5689655 + 0.7586207*42
```

## Part D

Plotting the regression line in 'red' can be done using the below code,

```{r plotting}
plot(icecreamdata$temperature,icecreamdata$sales,main="Sales VS Temperature",
     xlab = "Temperature",ylab = "Sales")
abline(lsfit(icecreamdata$temperature,icecreamdata$sales)$coefficients,
       col="blue")
```

For changing the regression line color,

```{r plottingred}
plot(icecreamdata$temperature,icecreamdata$sales,main="Sales VS Temperature",
     xlab = "Temperature",ylab = "Sales")
abline(lsfit(icecreamdata$temperature,icecreamdata$sales)$coefficients,
       col="red")
```
